Mapping algorithms to hardware\\ the AHIR story

### 2015-02-11 Documents/Presentations/eeColloquium/Talk.tex
Added ee colloquium talk.

\documentclass{beamer}
\usepackage{epsfig}
\title{IV EURASIP Seminar on the Hardware Design of DSP Systems\\
Program-to-hardware transformation and its use in heterogenous high-performance DSP}
\author{Madhav P. Desai \\ Department of Electrical Engineering\\
  Indian Institute of Technology -- Bombay, Powai, Mumbai -- 400076, INDIA}
\date{March 6, 2013}
\begin{document}

@@

\frame[containsverbatim]{\frametitle{Overview}

High-performance DSP can take advantage of several platforms:
\vspace{0.1in}

\begin{itemize}
\item Processors such as multi-cores, GPUs: ``easy'' to use, but power
	dissipation is an issue.
\item Programmable hardware:  lots of parallelism in principle, potential
	power benefits, difficult to map algorithms to  platform.  Can they
	beat power-performance characteristics of processors?
\item ASICs:  lots of parallelism, high efficiency, will surely beat power-performance
	characteristics of processors.  Difficult to design, and expensive to
	manufacture.
\end{itemize}

In terms of efficiency (performance/watt):
\begin{verbatim}
ASIC  > >  Processor, FPGA
Processor < ? >  FPGA 
\end{verbatim}

@@

@@

\frame[containsverbatim]{\frametitle{A system in AHIR}

@@

\begin{itemize}
\item A system description consists of a collection of modules (a module is
similar to a function in a C program).  The modules communicate
with the environment, with each other (through calls, and through
FIFO pipes), and with memory subsystems. 
\end{itemize}

@@

@@

\frame[containsverbatim]{\frametitle{The storage subsystem ${\bf S}$}
\begin{itemize}
\item The storage subsystem connects to the load/store operators in the datapath,
and is required to eventually service the load/store requests.
\item Declared variables in the source program are grouped
into storage spaces.  
\item The storage spaces are determined by
static reference analysis (the system can have
several distinct memory subsystems).
\item The memory subsystems have a relaxed consistency model: all accesses are time-stamped,
and accesses to the same memory location are finished in first-come-first-served
order.
\end{itemize}
}

@@

\frame[containsverbatim]{\frametitle{Dot-product circuit} 
\begin{figure}
\begin{centering}
\centerline{\psfig{figure=dotP.eps,width=4.0in,height=3.0in}}
\end{centering}
\end{figure}
}

@@

\frame[containsverbatim]{\frametitle{Future directions for our research}
\begin{itemize}
\item Applications: signal processing (image/multi-media), cryptology, 
CFD, Databases etc.
\item More optimizations:  latency reduction in non-parallel paths.
\item Algorithm-to-ASIC.
\item OpenMP, OpenCL support.
\end{itemize}
}

@@

The AHIR flow is open-source.  If you wish to try
it out, please get in touch with me: madhav@ee.iitb.ac.in

>>>>

\documentclass{beamer}
\usepackage{epsfig}
\title{Mapping algorithms to hardware\\ the AHIR story}
\author{Madhav P. Desai \\ Department of Electrical Engineering\\
  Indian Institute of Technology -- Bombay, Powai, Mumbai -- 400076, INDIA}
\date{January 21, 2015}
\begin{document}

@@

\frame[containsverbatim]{\frametitle{The premise}

\begin{itemize}
\item Hardware capacity has scaled, but design capability has remained stagnant.
\begin{itemize}
\item evidence: anecdotal, most chips today are cut-and-paste/repetitive.
\end{itemize}
\item Hardware capability/efficiency is not fully exploited in implementing complex algorithms.
\end{itemize}

@@

\frame[containsverbatim]{\frametitle{Acknowledgements}

Sameer Sahasrabudhe worked out many of the basic ideas as part of
his Ph.D. thesis  and also contributed most of the initial code for the first version of the tools. 
Thanks to Kavi Arya for many interesting discussions and inputs.  Maybe some day we will have
a front end for functional programming languages.
Kartik Lakhotia contributed to the FPGA implementation effort.
Thanks to Teemu Rinta-Aho and G. Caffarena for providing valuable user feedback.
}

@@

\frame[containsverbatim]{\frametitle{The starting point: the algorithm}
We view the algorithm as a program:
\begin{itemize}
\item which describes the behaviour of a collection of static threads (no spawning on the fly).
\item the threads interact using shared memory and FIFOs.
\item can be expressed in C.
\end{itemize}
}

\frame[containsverbatim]{\frametitle{The end point: a logic circuit}

@@

\begin{itemize}
\item A collection of modules (a module is
similar to a function in a C program).  The modules communicate
with the environment, with each other (through calls, and through
FIFO pipes), and with memory subsystems. 
\end{itemize}

@@

\frame[containsverbatim]{\frametitle{Type-2 petri-nets} 
Type-2 petri-nets\footnote{Introduced in Sameer Sahasrabudhe's Ph.D. thesis} 
are live and safe petri-nets constructed using a
set of simple rules.  The simplest type-2 petri-net has the form
shown in the figure.
\begin{figure}
\begin{centering}
\centerline{\psfig{figure=Type2Base.eps,width=2.5in,height=2.0in}}
\end{centering}
\end{figure}
}

\frame[containsverbatim]{\frametitle{Production rules} 
Replace a transition by a transition-place-transition (any region produced
from a transition is called a t-region):
\begin{figure}
\begin{centering}
\centerline{\psfig{figure=Type2T1.eps,width=2.5in,height=1.5in}}
\end{centering}
\end{figure}
}

\frame[containsverbatim]{\frametitle{Production rules} 
Replace a transition by an acyclic graph constructed using
transitions and p-regions: the acyclic graph must have a
single source and a single sink.
\begin{figure}
\begin{centering}
\centerline{\psfig{figure=Type2T2.eps,width=2.5in,height=2.0in}}
\end{centering}
\end{figure}
}

\frame[containsverbatim]{\frametitle{Production rules} 
Replace a place by a possibly cyclic graph consisting of
places and t-regions.
\begin{figure}
\begin{centering}
\centerline{\psfig{figure=Type2P1.eps,width=2.3in,height=2.0in}}
\end{centering}
\end{figure}
}

\frame[containsverbatim]{\frametitle{Production rules: pipelines} 
Using token-preserving regions, we can construct a well-behaved
pipeline.
\begin{figure}
\begin{centering}
\centerline{\psfig{figure=PipelineTPR.eps,width=4.5in,height=2.0in}}
\end{centering}
\end{figure}

Note: this can be generalized to DAG.
}

\frame[containsverbatim]{\frametitle{Liveness and Safety}

\begin{itemize}
\item   Assume that an enabled input transition eventually fires.  Then the type-2
petri-net is live and safe.
\item The type-2 petri-net is also powerful enough to describe the control
flow in imperative languages such as C/C++ as well as in synchronous
languages such as Esterel.
\end{itemize}

@@

\frame[containsverbatim]{\frametitle{CP-DP interactions}
\begin{figure}
\begin{centering}
\centerline{\psfig{figure=CPxDPxInteraction.eps,width=4.0in,height=3.0in}}
\end{centering}
\end{figure}
}

\frame[containsverbatim]{\frametitle{The storage subsystem ${\bf S}$}
\begin{itemize}
\item The storage subsystem connects to the load/store operators in the datapath,
and is required to eventually service the load/store requests.
\item Declared variables in the source program are grouped
into storage spaces.  
\item The storage spaces are determined by
static reference analysis (the system can have
several distinct memory subsystems).
\item The memory subsystems impose a strong consistency model: all accesses are time-stamped,
and accesses to the same memory location are finished in first-come-first-served
order.
\end{itemize}
}

@@

\frame[containsverbatim]{\frametitle{Dot-product virtual circuit} 
\begin{figure}
\begin{centering}
\centerline{\psfig{figure=dotP.eps,width=4.0in,height=3.0in}}
\end{centering}
\end{figure}
}

\frame[containsverbatim]{\frametitle{Optimizations} 
\begin{itemize}
\item Local dependency analysis: extract local parallelism.
\item Operator sharing: e.g. non-concurrent operations can be mapped to the same
physical operator.
\item Memory decomposition: different data structures can be mapped to different
memories.
\item Loop pipelining: the hardware can keep multiple iterations of a loop in
progress so that throughput is improved.
\end{itemize}
}

\frame[containsverbatim]{\frametitle{Memory decomposition: static aliasing analysis}
\begin{verbatim}
int A, B, C, D; // storage objects.
int foo()
{
   int ret_val = 0;
   int* p1 = &A;
   ret_val += *p1;
   int *p2 = &B;
   ret_val += *p2;
   int *p3 = &C;
   ret_val += *p3;
   p3 = &D;
   ret_val += *p3;
   return(ret_val);
}
\end{verbatim}
}

\frame[containsverbatim]{\frametitle{Memory decomposition: grouping}

\begin{itemize}
\item The memory spaces are $\{A\}$, $\{ B\}$, $\{ C, D \}$.
\item $C$ and $D$ are grouped together because pointer $p3$ can
point to either $C$ or $D$.
\end{itemize}

\frame[containsverbatim]{\frametitle{Loop pipelining}
\begin{verbatim}
void Daemon()
{
  while(1)
  {
    float a = read_float32("a_pipe");
    float b = read_float32("b_pipe");
    float c = read_float32("c_pipe");
    write_float32("d_pipe", (a+b)*c);
  }
}
\end{verbatim}
}

\frame[containsverbatim]{\frametitle{Loop Pipelining: dependencies across iterations}
\begin{itemize}
\item Loop pipelining involves executing multiple iterations of a
loop at the same time.
\item While doing this, all dependencies must be taken care of.
\begin{itemize}
\item Operation order.
\item Data dependency.
\item Memory accesses.
\item Pipe accesses.
\end{itemize}
\end{itemize}
Currently, we are a bit conservative.  High performance compiler
community has made substantial progress here.
}

@@

\frame[containsverbatim]{\frametitle{Future directions for our research}
\begin{itemize}
\item Applications: signal processing (image/multi-media), cryptology, 
CFD, Databases etc.
\item Optimizations:  latency reductions, alternative DP structures.
\item Algorithm-to-ASIC, potential asynchronous implementations.
\item Front-end stuff:
\begin{itemize}
\item Other front-ends (going beyond C), OpenMP, OpenCL support.
\item More aggressive compiler optimizations.
\end{itemize}
\end{itemize}
}

@@

The AHIR flow is open-source.  If you wish to try
it out, please get in touch with me: madhav@ee.iitb.ac.in
or search on github (standard disclaimers apply).
